Sparse Function-space Representation \\ of Neural Networks

### 2023-05-17 paper/main.tex
Link colours
reworked uci

% CL section
%\textbf{Function-space methods for sequential learning} The central problem in continual learning is how to deal with the non-stationary nature of the training distribution, which causes the training process to overwrite the previously learnt parameters---leading to the model forgetting the previously learnt functions. 
%The approaches proposed in the CL literature can be categorized into inference-based, rehearsal-based, and model-based methods; we refer the reader to \citep{parisi2019continual, de2021continual} for complete reviews. The methods in the first category usually tackle this problem with weight-space regularization, such as EWC~\citep{kirkpatrick2017overcoming}, SI~\citep{zenke2017continual}), or VCL~\citep{nguyen-tuongModel2009} that induce retention of previously learned information in the weights alone.
%However, encoding functional knowledge in the weights does not guarantee a good quality of the predictions, thus a better to introduce a function-space regularization \citep{li2018lwf, benjamin2018measuring, titsias2019functional, buzzega2020dark, pan2020continual, rudner2022continual}. These methods bridge the gap between objective and rehearsal-based methods since they rely on a subset of the training data for each task to compute the regularization term. Recently proposed methods, such as DER~\citep{buzzega2020dark}, FROMP~\citep{pan2020continual}, and S-FSVI~\citep{rudner2022continual} achieve state-of-the-art performances among the objective-based techniques in several CL benchmarks.

\textbf{Function-space methods for sequential learning} 
Continual learning (CL) hinges on how to deal with a non-stationary training distribution, wherein the training process may overwrite previously learned parameters, causing catastrophic forgetting \citep{mccloskey1989catastrophic}. CL approaches fall into inference-based, rehearsal-based, and model-based methods (see \cite{parisi2019continual, de2021continual} for an overview). The methods in the first category usually tackle this problem with weight-space regularization methods, such as EWC~\citep{kirkpatrick2017overcoming}, SI~\citep{zenke2017continual}), or VCL~\citep{nguyen-tuongModel2009} that induce retention of previously learned information in the weights alone. However, regularizing the weights does not guarantee good quality predictions. Function-space regularization techniques~\cite{li2018lwf, benjamin2018measuring, titsias2019functional, buzzega2020dark, pan2020continual, rudner2022continual} address this by relying on training data subsets for each task to compute a regularization term. These methods bridge the gap between objective and rehearsal-based methods \todo{I'm not familiar with the differences between objective vs rehearsal based methods, and it isn't explained here} since they use a subset of the training data for each task to compute the regularization term. Recently proposed methods, \eg, DER~\citep{buzzega2020dark}, FROMP~\citep{pan2020continual}, and S-FSVI~\citep{rudner2022continual} achieve state-of-the-art performance among the objective-based techniques in several CL benchmarks. \todo{Are DER, FROMP and FSVI function-space regularization methods? I assume so but the text makes it a bit vague.}\looseness-2

@@

\subsection{Capturing uncertainty in UCI tasks under supervised learning}
\label{sec:uci}
%
We consider eight UCI benchmarks~\cite{UCI}, covering of a variety of binary and multi-class classification tasks with different data set sizes. We initially train a two-layer MLP for each of the classification tasks and follow the setup in \cref{sec:methods} for \our. That is, after training we construct the \our dual and use the resulting model for uncertainty quantification. For comparison, we used the same hyperparameters for the neural network training as in the UCI experiments of \cite{immer2021improving}, and similarly perform hyperparameter search over the prior precision $\delta$, and run the experiment over $10$ random splits. In \cref{tbl:uci} (left), we compare our model to the NN MAP estimate, mean-field VI, a Bayesian NN, and a GLM model (see \cite{immer2021improving} for the baselines) in negative log-predictive density (NLPD). We match the predictive performance despite being sparse (here, $M=256$). Results for $M=16,32,64,128,256$ are included in \cref{app:uci} with full experiment details.

In \cref{tbl:uci} (right), we also include an ablation and comparison study (to \cite{immer2021improving}) with a fixed low number of inducing points ($M=32$). The comparison shows that our sparse method is able to summarize the information of the full data set more efficiently than sparse subset methods, and can match the performance of the full GP even with a low number of inducing points. This is even more apparent from \cref{fig:uci} which shows the benefits of \our in summarizing the full data distribution onto a small set of inducing points over just picking a random subset.

@@

% Discussion points
In practical terms, our model serves a role similar to a sparse GP. However, unlike a conventional GP, it does not provide a straightforward method for specifying or tuning the prior covariance function. This limitation can be addressed indirectly: the architecture of the neural network and the choice of activation functions can be used to implicitly specify and tune the prior assumptions, thereby incorporating a broad range of inductive biases into the model. It is important to note that the NTK linearizes the network around the MAP solution $\weights^{*}$, resulting in the function-space prior (and consequently the posterior) being only a locally linear approximation of the neural network model.

%In contrast to conventional GP kernels, the NTK does not have any hyperparameters which need to be learned.
%On the other hand, the NTK may be highly non-stationary and is dependent on the NN architecture, for example, the activation functions.

The broader impact of this work lies in its potential to provide tooling for how neural networks are utilized, offering more efficient and principled ways of handling uncertainty and continual learning. This contribution, we believe, has significant implications for future applications of machine learning in dynamic, real-world settings where data is unevenly distributed, uncertain, and continuously evolves.\looseness-1

>>>>

% CL section
%\textbf{Function-space methods for sequential learning} The central problem in continual learning is how to deal with the non-stationary nature of the training distribution, which causes the training process to overwrite the previously learnt parameters---leading to the model forgetting the previously learnt functions. 
%The approaches proposed in the CL literature can be categorized into inference-based, rehearsal-based, and model-based methods; we refer the reader to \citep{parisi2019continual, de2021continual} for complete reviews. The methods in the first category usually tackle this problem with weight-space regularization, such as EWC~\citep{kirkpatrick2017overcoming}, SI~\citep{zenke17a}), or VCL~\citep{nguyen-tuongModel2009} that induce retention of previously learned information in the weights alone.
%However, encoding functional knowledge in the weights does not guarantee a good quality of the predictions, thus a better to introduce a function-space regularization \citep{li2018lwf, benjamin2018measuring, titsias2019functional, buzzega2020dark, pan2020continual, rudner2022continual}. These methods bridge the gap between objective and rehearsal-based methods since they rely on a subset of the training data for each task to compute the regularization term. Recently proposed methods, such as DER~\citep{buzzega2020dark}, FROMP~\citep{pan2020continual}, and S-FSVI~\citep{rudner2022continual} achieve state-of-the-art performances among the objective-based techniques in several CL benchmarks.

\textbf{Function-space methods for sequential learning} 
Continual learning (CL) hinges on how to deal with a non-stationary training distribution, wherein the training process may overwrite previously learned parameters, causing catastrophic forgetting \citep{mccloskey1989catastrophic}. CL approaches fall into inference-based, rehearsal-based, and model-based methods (see \cite{parisi2019continual, de2021continual} for an overview). The methods in the first category usually tackle this problem with weight-space regularization, such as EWC~\citep{kirkpatrick2017overcoming}, SI~\citep{zenke17a}), or VCL~\citep{nguyen-tuongModel2009} that induce retention of previously learned information in the weights alone. However, regularizing the weights does not guarantee good quality predictions. Function-space regularization techniques~\cite{li2018lwf, benjamin2018measuring, titsias2019functional, buzzega2020dark, pan2020continual, rudner2022continual} address this by relying on training data subsets for each task to compute a regularization term. These methods bridge the gap between objective and rehearsal-based methods \todo{I'm not familiar with the differences between objective vs rehearsal based methods, and it isn't explained here} since they use a subset of the training data for each task to compute the regularization term. Recently proposed methods, \eg, DER~\citep{buzzega2020dark}, FROMP~\citep{pan2020continual}, and S-FSVI~\citep{rudner2022continual} achieve state-of-the-art performance among the objective-based techniques in several CL benchmarks. \todo{Are DER, FROMP and FSVI function-space regularization methods? I assume so but the text makes it a bit vague.}\looseness-2

@@

\subsection{Capturing uncertainty in UCI tasks under supervised learning}
\label{sec:uci}
%
We first evaluate \our on eight UCI benchmarks~\cite{UCI}, a variety of binary and multi-class classification tasks with different data set sizes.
We train a two-layer MLP for each of the classification tasks and follow the experiment set up in \cite{immer2021improving}.
%by using the same hyperparameters, performing a hyperparameter search over the prior precision $\delta$,
% We follow \cite{immer2021improving} by using the same hyperparameters, performing a hyperparameter search over the prior precision $\delta$,
% and run the experiment over $10$ random splits.
% That is, after training we construct the \our dual and use the resulting model for uncertainty quantification.
% for the neural network training as in the UCI experiments of
\cref{tbl:uci} (left) shows that \our with $M=256$ either matches or outperforms the predictive performance of the NN MAP, mean-field VI, Bayesian NN, and GLM
predictions %in terms of negative log-predictive density (NLPD).
(baselines from \cite{immer2021improving}).
That is, we match predictive performance despite being sparse.

In \cref{tbl:uci} (right) we compare to the subset GP method from \cite{immer2021improving} whilst using only $M=32$ inducing points.
It shows that \our is able to summarize the full data set more effectively than the GP subset method as it maintains predictive performance
whilst using fewer inducing points.
\cref{fig:uci} further shows that as the number of inducing points is lowered $M=512\rigtharrow 16$, \our is able to maintain a much better NLPD than the GP subset.
These results demonstrate \our's sparse representation captures information from the entire data set and as a result provides good uncertainty estimates.
We provide full details of our experiments and further results using $M=16,32,64,128,256$ in \cref{app:uci}.
% benefits of \our in summarizing the full data distribution onto a small set of inducing
% points over just picking a random subset.

% We match the predictive performance despite being sparse (here, $M=256$).

% This is
% However, our method used

% despite being sparse (here, $M=256$).

% (left), we compare our model to the NN MAP estimate, mean-field VI, a Bayesian NN, and a GLM model (see \cite{immer2021improving} for the baselines) in negative log-predictive density (NLPD).
% In \cref{tbl:uci} (left), we compare our model to the NN MAP estimate, mean-field VI, a Bayesian NN, and a GLM model (see \cite{immer2021improving} for the baselines) in negative log-predictive density (NLPD).
% We match the predictive performance despite being sparse (here, $M=256$).
% Results for $M=16,32,64,128,256$ are included in \cref{app:uci} with full experiment details.

% \cref{tbl:uci} (right) shows that \our is able to summarize the information of the full data set more efficiently than the sparse GP subset method.
% As a result, it maintains predictive performance whilst using fewer inducing points.
% This is even more apparent from \cref{fig:uci} which shows the benefits of \our in summarizing the full data distribution onto a small set of inducing points over just picking a random subset.

% In \cref{tbl:uci} (right), we also include an ablation and comparison study (to \cite{immer2021improving}) with a fixed low number of inducing points ($M=32$).
% The comparison shows that our sparse method is able to summarize the information of the full data set more efficiently than sparse subset methods, and can match the performance of the full GP even with a low number of inducing points. This is even more apparent from \cref{fig:uci} which shows the benefits of \our in summarizing the full data distribution onto a small set of inducing points over just picking a random subset.

@@

% Discussion points
It is worth noting that the NTK linearises the network around the MAP solution $\weights^{*}$,
so the function-space prior (and thus posterior) is only a locally linear approximation.
In contrast to conventional GP kernels, the NTK does not have any hyperparameters which need to be learned.
On the other hand, the NTK may be highly non-stationary and is dependent on the NN architecture, for example, the activation functions.

The broader impact of this work lies in its potential to provide tooling for how neural networks are utilized, offering more efficient and principled ways of handling uncertainty and continual learning. This contribution, we believe, has significant implications for future applications of machine learning in dynamic, real-world settings where data is unevenly distributed, uncertain, and continuously evolves.
